---
title: "Problem Set 4"
author: "Ashley Knox"
date: "3/9/2017"
output:
  html_document: default
  (my html document keeps getting stuck when I try to knit it, but here's the code at least)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

# Submission

1. Fork this repository to your own account
1. Make changes to the `README.Rmd` file (including the author field above).
1. Knit the file to HTML format
1. Publish HTML on Rpubs (you will need to sign up for the free account) and *add the link below*.
1. Submit a PR with your answers.

*Link to Rpubs document:* (coming soon - Knit keeps getting stuck)

# Overview

Take the [datacamp course on joining data](https://www.datacamp.com/courses/joining-data-in-r-with-dplyr) to learn the join verbs.

You will analyze some data in the [`nihexporter`](https://github.com/jayhesselberth/nihexporter) package to answer some questions. The questions will cover methods in `dplyr` including two-table verbs.

First you need to install the `nihexporter` package (and `devtools`, if you don't have it already).

```r
install.packages('devtools')
devtools::install_github("jayhesselberth/nihexporter")
```

Read the [Overview vignette](http://jayhesselberth.github.io/nihexporter/articles/nihexporter.html) to learn the structure of the pacakge. Take a look at two tables from the package:

```{r overview}
library(nihexporter)
library(tidyverse)
library(broom)

projects
project_io
```

# Problems

## Problem 1

The `projects` and `project_io` tables have the `r intersect(names(projects), names(project_io))` column in common. Use this column to perform a `left_join` and identify the most expensive project for each institute. Project institutes are in the `projects` table, whereas total project costs are available in the `project_io` table. Note you will have to use the `arrange` and `slice` verbs from `dplyr`.

```{r problem_1, eval=FALSE}

projects %>% select(institute, project.num, fy.cost) %>% left_join(., project_io) %>% select(project.num, institute, fy.cost, total.cost) %>% group_by(institute) %>% arrange(desc(total.cost)) %>% slice(1:1) %>% arrange(desc(total.cost)) -> total.cost_per_institute

total.cost_per_institute


```

## Problem 2

Create a plot that describes the change in output (number of publications) for the most expensive grant from the above list. Note that you cannot use `n.pubs` for this, as that is an aggregate count over time. You need a tbl of following structure, which you can then use to calculate the number of publications each year (hint: use `n()` in a `summarize` statement:

```{r p2_tbl}
#need to repeat code for table so knit works
projects %>% select(institute, project.num, fy.cost) %>% left_join(., project_io) %>% select(project.num, institute, fy.cost, total.cost) %>% group_by(institute) %>% arrange(desc(total.cost)) %>% slice(1:1) %>% arrange(desc(total.cost)) -> total.cost_per_institute
#projects and publinks have "project.num" in common
#projects+publinks and publications have "pmid"" in common
total.cost_per_institute %>% head(1) %>% left_join(publinks) %>% left_join(publications) %>% select(project.num, pub.year, fy.cost, total.cost, pmid) %>% na.omit() %>% select(project.num, institute, total.cost, pub.year, pmid) %>% group_by(project.num, institute, total.cost, pub.year) %>% summarize(total.pubs= n()) %>% na.omit() -> p2_tbl

p2_tbl
```

```{r problem_2}
ggplot(p2_tbl, aes(pub.year, total.pubs)) + geom_col(fill = "darkblue") + xlab('Publication Year') + ylab('Total Publications') -> total.pubs_vs_year

total.pubs_vs_year
```

## Problem 3

Use the `projects` and `clinical_studies` table to calculate how many clinical studies have been started by projects from each institute.

```{r problem_3}
#projects and clinical_studies have project.num in common; join these, then select all but the "study.status" column. Group by institute and project.num so that summarise cna be used to find the total number of studies per project
select(projects, institute, project.num) %>% left_join(clinical_studies) %>% na.omit() %>% select(institute, project.num, trial.id) %>% group_by(institute, project.num) %>% summarise(total.studies = n()) %>% arrange(desc(total.studies)) -> p3_tbl

p3_tbl
```

## Problem 4

Calculate correlations between `total.cost` and `n.pubs` for each institute. You will need to `join` the `projects` and `project_io` tables to link `institute` with `total.cost` and `n.pubs`. You will then need to use `group_by` and `do` from `dplyr` and `tidy` from `broom` to calculate correlations.

```{r problem_4, eval=FALSE}
projects %>% left_join(project_io) %>% select(institute, total.cost, n.pubs) %>% na.omit() %>% group_by(institute) %>% summarize(total.cost = sum(total.cost), total.pubs = sum(n.pubs)) %>% do(tidy(cor(.$total.cost, .$total.pubs))) -> cost.pubs.coeff

cost.pubs.coeff

```

## Problem 5 (extra credit)

Correlate [`rcr`](https://icite.od.nih.gov/) values from the `publications` table to one of:

- total cost (overall for NIH, or by institute)
- budget mechanism (e.g., RFA, PAR, etc.)
- type of grant (e.g. R01, P01, R21, etc.). 

```{r extra_credit}
left_join(publications, publinks) %>% left_join(project_io) %>% na.omit() %>% select(pmid, project.num, rcr, total.cost) %>% group_by(pmid) %>% do(tidy(cor(.$total.cost, .$rcr)))
```

## Problem 6 (extra extra credit)

Use [`gganimate`](https://github.com/dgrtwo/gganimate) to illustrate the change in cost of R01s for a few institutes over time.

```{r extra_extra_credit}
```
